\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{anysize, longtable}
\usepackage{amsthm, amsfonts, amsmath, amssymb, booktabs, float}
\usepackage{sectsty, csquotes, xcolor, xspace, setspace, listings}
\usepackage{graphicx, tabularx, changepage, multirow, subcaption}
\usepackage{algorithm, algpseudocode, multicol, cuted, colortbl}
\usepackage{makecell, supertabular}
\usepackage[hidelinks,bookmarksdepth=3]{hyperref}
\usepackage[capitalise]{cleveref}
\usepackage[inline]{enumitem}
\usepackage[english]{babel}
\usepackage{microtype}
\usepackage{libertinus}
\usepackage[scale = 0.85]{plex-mono}
\usepackage[maxbibnames=99]{biblatex}
\usepackage{titlesec}
\usepackage{tikz}

\usetikzlibrary{fit, shapes.geometric, shapes.symbols, positioning, shapes.misc, calc, arrows.meta, patterns, backgrounds, decorations, decorations.markings, decorations.pathreplacing, shadows, hobby, decorations.pathmorphing}

\makeatletter

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}

\theoremstyle{remark}
\newtheorem*{note}{Note}
\newtheorem*{remark}{Remark}

\newcommand{\inlinecode}[1]{\lstinline[basicstyle=\ttfamily,columns=fixed]$#1$}
\lstset{basicstyle=\setstretch{0}\small\ttfamily,breaklines=true,columns=fixed,captionpos=b,xleftmargin=0.5cm}

\newlength{\offsetpage}
\newenvironment{widepage}[1][2cm]{
  \setlength{\offsetpage}{#1}
  \begin{adjustwidth}{-\offsetpage}{-\offsetpage}
  \addtolength{\linewidth}{2\offsetpage}
}{\end{adjustwidth}}

\algnewcommand{\IIf}[1]{\State \algorithmicif\ #1\ \algorithmicthen}
\algnewcommand{\IElse}[1]{\algorithmicelse\ #1}
\algnewcommand{\EndIIf}{\unskip\ \algorithmicend\ \algorithmicif}

\addbibresource{biblio.bib}

\allowdisplaybreaks
\marginsize{2.0cm}{2.0cm}{1.25cm}{1.25cm}
\setcounter{tocdepth}{2}
\setlist[itemize]{topsep=0pt, partopsep=0pt, parsep=0pt, itemsep=\parskip}
\setlist[enumerate]{topsep=0pt, partopsep=0pt, parsep=0pt, itemsep=\parskip}
\titlespacing*{\paragraph}{0pt}{0.5\baselineskip}{0.5\baselineskip}
\captionsetup[figure]{aboveskip=5pt}
\captionsetup[subfigure]{aboveskip=3pt}

\title{Real-Time Analysis of Public GitHub Activity}
\author{Roland Bernard, 19598, \inlinecode{rolbernard@unibz.it}}
\date{February 2026}

\def\maketitle{
\begin{strip}
  \begin{adjustwidth}{5mm}{5mm}
    \vspace{-4.5em}
    \noindent{\small\bfseries Project Report -- Real-Time Big Data Processing 2024/2025\par}
    \vskip 0.25em
    \noindent{\LARGE\bfseries \@title\par}
    \vskip 1em
    \noindent{\large \@author\par}
    \vskip 1em
    \noindent{\small \@date\par}
    \vskip 1.5em
    \noindent{
      \small
      \begin{tabular}{rl}
        Code Repository & \url{https://gitlab.inf.unibz.it/Roland.Bernard/rtbdp-project}, or \\
                        & \url{https://github.com/rolandbernard/rtbdp-project} \\
        Commit & TODO (latest commit in the repository) \\
        Live Demo Link & \url{https://rtgh.rolandb.com/} \\
      \end{tabular}
    }
    \vskip 0.5em
  \end{adjustwidth}
\end{strip}
}
\renewenvironment{abstract}{
  \vskip 1em
  \begin{quote}
  {\normalsize \abstractname\par}
  \small
}{
  \end{quote}
  \vskip 1em
}

\makeatother

\input{tikzfigures/defs.tex}

\begin{document}

\maketitle

\begin{abstract}
The open-source software ecosystem generates large volumes of data. Much of this activity takes place on GitHub. While native and third-party analytics tools exist, they predominantly focus on historical data or static snapshots. This project addresses that gap by developing a real-time big data processing application for analyzing public GitHub activity.

The implemented system utilizes a pipeline orchestrated via Docker Compose, leveraging Apache Kafka for messaging and data ingestion from the GitHub Events API. Processing is handled by Apache Flink, which performs complex stream processing tasks including data enrichment, windowed aggregations, and leaderboard generation.

The final application features an interactive single-page application (SPA) frontend built with React and TypeScript, which visualizes data through live event feeds, dynamic charts, and real-time leaderboards. The system successfully processes a velocity of 70-120 events per second, handling challenges such as out-of-order event delivery and real-time ranking updates.
\end{abstract}

\input{1.domain.tex}

\input{2.dataset.tex}

\input{3.architecture.tex}

\input{4.functionalities.tex}

\input{5.conclusion.tex}

\printbibliography

\end{document}
